---
title       : Child Height Prediction
subtitle    : using Galton data
author      : Phi Primed
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---
```{r echo=FALSE}
library(UsingR)
data(galton)
```
## Synopsis
- A web app that predicts child's height based on the input father's and mother's height
- Uses the _galton_ dataset from the _UsingR_ package
    - built from a tabulated data set used by Galton in 1885 
        - relationship between a parent's height and their children's
    - 928 observations on 2 variables (928 children and 205 parents)
        - child : the child's height
        - parent : the "midparent" height 
            - average of the fathers height and 1.08 times the mother's
- A linear model is built from the dataset for the purpose of child height prediction
- The predicted height is shown along with the 95% confidence interval
- Frequency distributions of parent's height and child height are also shown
    - purpose is to illustrate _Regression toward the mean_

--- .class #id 

## Providing Input
- The mean, min and max values for parent height in _galton_
```{r echo=FALSE}
paste("Mean: ",mean(galton$parent))
paste("Min: ",min(galton$parent))
paste("Max: ",max(galton$parent))
```

- Since the above values are for "midparent" height, inputs slightly lower and higher are allowed for father's and mother's height
- Use the sliders to provide input values for father's height and mother's height
    - initial defaults are set to be close to the mean
- "midparent" height is computed by the app using your inputs and displayed

---
## Prediction using a Linear Model
- The linear model built using the _galton_ dataset

```{r smallplot, fig.width=5, fig.height=4, echo=FALSE }
model <- lm(child ~ parent, data=galton)
plot(galton$parent, galton$child, xlab="Midparent height(inches)",
           ylab="Child's height (inches)",
           main="Galton data set: Linear Regression fit")
lines(galton$parent, model$fitted, col="red", lwd=3)
```
- A sample prediction (fit - fitted value lwr,upr - bounds of 95% confidence interval)
```{r collapse=TRUE}
predict(model, data.frame(parent=c(70)), interval="confidence")
```

---
## Regression towards the mean
- When "midparent" height is significantly higher than the mean, child height is not as much higher than its mean
- When "midparent" height is significantly lower than the mean, child height is not as much lower than its mean
- This is illustrated via the frequency distribution plots displayed by the app
